Causal analysis of COVID-19 observational data in German districts reveals effects of mobility, awareness, and temperature

### 2020-07-28 manuscript/v2_medrxiv/results.Rmd
small fixes (range of exposure variables)

"Awareness" had two opposite effects on the outcome in our DAG. Awareness measured by Google searches for *corona* had a positive effect on the number of reported cases. An one percent point increase of the state's Google searches (relative to other states and the observation period) leads to an increase of approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Searches corona")%>%pull("effect estimate")), 1)` percent. For example, if a district shows 10 percent points more relative searches for *corona* than another one, we expect approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Searches corona")%>%pull("effect estimate"))^(10)), 0)` percent more infections for this district after 8 days.  _COVID-19 burden_ (reported number of cases on day of exposure) affected the outcome negatively, where every additional daily case in the district leads to a `r round(100*abs(1-causaleffects%>%filter(cause=="COVID-19 burden")%>%pull("effect estimate")), 1)` percent decrease in newly reported case numbers. The last plot in Figure \@ref(fig:effectsrange) visualizes this relationship: For a local outbreak with 25 daily cases as COVID-19 burden, we estimate as total causal effect a subsequent reduction of infection rate by `r round(100*abs(1-(causaleffects%>%filter(cause=="COVID-19 burden")%>%pull("effect estimate"))^(25-mymeans["COVID-19 burden"])), 1)` percent.

Within our model, we observed a causal effect of temperature and all other weather variables. Every increase of 1 degree Celsius in temperature leads to a reduction of the daily reported case numbers by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Temperature")%>%pull("effect estimate")), 1)` percent. On the other hand, we found an increasing effect of rainfall: One millimeter (=1 liter per square meter) more rainfall leads to an increase of reported case numbers by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Rainfall")%>%pull("effect estimate")), 1)` percent. We observe effects for humidity and wind as well (higher humidity leading to more cases, stronger wind leading to less cases). In perspective (Figure \@ref(fig:effectsrange)), with temperature we expect an increase by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Temperature")%>%pull("effect estimate"))^(0-mymeans["Weather (temperature)"])), 1)` percent at a daily average temperature of $0^{\circ}C$. For rainfall, we expect on a rainy day with 10 mm rainfall a corresponding increase of the infection rate by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Rainfall")%>%pull("effect estimate"))^(10-mymeans["Weather (rainfall)"])), 1)` percent.

>>>>

"Awareness" had two opposite effects on the outcome in our DAG. Awareness measured by Google searches for *corona* had a positive effect on the number of reported cases. An one percent point increase of the state's Google searches (relative to other states and the observation period) leads to an increase of approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Searches corona")%>%pull("effect estimate")), 1)` percent. For example, if a district shows 10 percent points more relative searches for *corona* than another one, we expect approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Searches corona")%>%pull("effect estimate"))^(10)), 0)` percent more infections for this district after 8 days.  _COVID-19 burden_ (reported number of cases on day of exposure) affected the outcome negatively, where every additional daily case in the district leads to a `r round(100*abs(1-causaleffects%>%filter(cause=="COVID-19 burden")%>%pull("effect estimate")), 1)` percent decrease in newly reported case numbers. The last plot in Figure \@ref(fig:effectsrange) visualizes this relationship: For a local outbreak with 20 daily cases as COVID-19 burden, we estimate as total causal effect a subsequent reduction of infection rate by `r round(100*abs(1-(causaleffects%>%filter(cause=="COVID-19 burden")%>%pull("effect estimate"))^(20-0)), 1)` percent.

Within our model, we observed a causal effect of temperature and all other weather variables. Every increase of 1 degree Celsius in temperature leads to a reduction of the daily reported case numbers by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Temperature")%>%pull("effect estimate")), 1)` percent. On the other hand, we found an increasing effect of rainfall: One millimeter (=1 liter per square meter) more rainfall leads to an increase of reported case numbers by approximately `r round(100*abs(1-causaleffects%>%filter(cause=="Rainfall")%>%pull("effect estimate")), 1)` percent. We observe effects for humidity and wind as well (higher humidity leading to more cases, stronger wind leading to less cases). In perspective (Figure \@ref(fig:effectsrange)), with temperature we expect an increase by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Temperature")%>%pull("effect estimate"))^(0-20)), 1)` percent at a daily average temperature of $0^{\circ}C$ compared to a day with $20^{\circ}C$. For rainfall, we expect on a rainy day with 10 mm rainfall a corresponding increase of the infection rate by approximately `r round(100*abs(1-(causaleffects%>%filter(cause=="Rainfall")%>%pull("effect estimate"))^(10-0)), 1)` percent compared to a day with no precipitation.
